---
title: "p8105_hw6_yq2378"
author: "Qi Yumeng"
date: "2023-12-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library and seed, warning = FALSE}
library(tidyverse)
library(plotly)
library(ggpubr)
set.seed(1)
```
# Problem 1

```{r load homicide data}
homicide = read.csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")

homicide = homicide |>
  mutate(city_state = str_c(city, state, sep=", "),
         if_solved = if_else(disposition == "Closed by arrest",1,0 ),
         victim_age = as.numeric(victim_age)) |>
  filter(!(city_state %in% c('Dallas, TX', 'Phoenix, AZ', 'Kansas City, MO','Tulsa, AL'))) |>
  filter(victim_race %in% c("White", "Black"),
         !is.na(victim_age),
         victim_sex %in% c("Female", "Male")) # delete non-numeric, missing age and Unknown sex
```

```{r build a function}
# Function to fit logistic regression and extract OR and CI
fit_logistic = function(data) {
  glm_res = glm(if_solved ~ victim_age + victim_sex + victim_race, data = data, family = "binomial")
  tidy_res = broom::tidy(glm_res, exponentiate = TRUE, conf.int = TRUE)
  or_ci = tidy_res |>
    filter(term == "victim_sexMale") |>
    select(estimate, conf.low, conf.high)
  return(or_ci)
}
```

```{r Baltimore}
# Fit logistic regression for Baltimore, MD
homicide |>
  filter(city_state == "Baltimore, MD") |>
  fit_logistic()

```

```{r loop the city}
# Fit logistic regression for each city and extract OR and CI
city_results = homicide |>
  group_by(city_state) |>
  nest() |>
  mutate(or_ci = map(data, fit_logistic)) |>
  unnest(or_ci) |>
  select(-data) |>
  rename(OR = estimate, CI_low = conf.low, CI_high = conf.high) |>
  arrange(desc(OR))
```

```{r plot for ORs and CIs, fig.width = 15, fig.align = 'center'}
ggplot(city_results, aes(x = reorder(city_state, OR),
                         y = OR, ymin = CI_low, ymax = CI_high)) +
  geom_point(position = position_dodge(width = 0.5), size = 1, color = "red") +
  geom_errorbar(position = position_dodge(width = 0.5), width = 0.2) +
  geom_hline(yintercept = 1, color = "blue", linetype = 2) +
  labs(title = "Estimated ORs and CIs for Solving Homicides (Male vs. Female)",
       x = "City",
       y = "Odds Ratio") +
  theme_pubclean() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The plot provides a visual representation of the estimated odds ratios (ORs) and their corresponding confidence intervals (CIs) for 47 cities. Notably, New York, NY exhibits the smallest OR, while Albuquerque, NM boasts the largest.

Beginning with Atlanta, GA, a distinct trend emerges: the subsequent six cities exhibit ORs greater than 1. This implies that in these particular cities, homicide cases involving male victims are more likely to be resolved compared to those involving female victims. On the contrary, the remaining 41 cities in the dataset exhibit an inverse relationship, signifying that homicides with female victims tend to have higher odds of resolution.

Regarding the error bars in the plot, it's notable that the cities positioned towards the right side generally exhibit wider error bars. This observation suggests a potential issue related to sample size in these cities. Wider error bars typically indicate increased uncertainty in the estimated odds ratios, and this could be attributed to smaller sample sizes for certain cities.

# Problem 2


```{r load weather data}
weather = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

```

```{r beta prod}
# prepocessing? 

weather |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy)) |> 
  select(.id,results) |> 
  unnest(results) |> 
  filter(term %in% c('tmin', 'prcp')) |>
  group_by(.id) |>
  summarise(estimate = log(prod(estimate))) |>
  ggplot(aes(x = estimate)) + geom_density() + theme_pubclean()

```



```{r R2 prod}
# prepocessing? 
weather |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::glance)) |> 
  select(results) |> 
  unnest(results) |>
  select(r.squared) |>
  ggplot(aes(x = r.squared)) + geom_density() + theme_pubclean()

```